Real Time Variational Density Forecasts

### 2017-02-03 Confirmation/confirmation_V2.tex
electricity section started

Electricity prices have been subject to an extensive forecasting literature, as detailed in \citet{Weron2014}, and can be extremely volatile. For the 2012-2015 period in Victoria, the mean price was X/KWH but on X occaisions the price spiked upwards several orders of magnitude to upwards of $1,000/KWH, reaching as high as $10,000/KWH on X date. For this reason market participants are increasingly interested in a predictive density forecast, as the probability that the price will rise above a particular threshold in the next time period is more important than the mean value. To calculate a suitably accurate predictive density the Bayesian methodology is to be adopted in this thesis, as it can average over parameter and model uncertainty that is often ignored in frequentist methods. See \citet{Geweke2006} for a review on Bayesian forecasting and \citet{Gneiting2014} for density forecasting in general.

@@

Alternatives to MCMC include Variational Bayes (VB) and Expectation Propogation (EP), which aim to substitute the true posterior distribution for a vector of unknown parameters $\theta$, denoted by $p(\theta | y)$, with some approximate distribution denoted by $q(\theta | \lambda)$. Here $\lambda$ is a set of auxillary parameters associated with the approximation and not the model. For example, if $q(\theta | \lambda)$ is Gaussian $\lambda$ would be the mean and variance that parameterise $q$. Variational Bayes will be explored as an alternative computation strategy to MCMC in the context of short term electricity load forecasting using data from the 2012-2015 period in Victoria. The advantages in computation time for VB will be compared to the loss of statisitcal accuracy associated with the use of an approximation of the posterior instead of the truth.

@@

Note: Will make the picture later
\begin{figure}
\label{MCMC}
%\includegraphics{MCMC.png}
\caption{MCMC draws for the AR(2) model with $T = 150, \phi_1 = 0.75, \phi_2 = 0.2$ and $\sigma^2 = 1$. Approximately 24\% of both $\phi$ parameters were accepted by the Metropolis Hastings step, then the first 50,000 draws were discarded and the remainder thinned to 1 in 300 to make dependence between draws negligible.}
\end{figure}

@@

Restricting the approximating distribution to a factorisable family may be unsatisfactory in many applications where the dependence between parameters is important, as in the AR(2) model where Figure (\ref{MCMC}) shows strong dependence between $\phi_1$ and $\phi_2$. Without factorisation the simple method to maximise the ELBO as in MFVB is unavailable. \citet{Paisley2012} and \citet{Ranganath2014} have adapted a gradient ascent algorithm for use in Variational Bayes, resulting in what is known as Stochastic Variational Bayes (SVB).

@@

Note: Don't have data on hand, will do this section shortly

>>>>

Electricity prices have been subject to an extensive forecasting literature and is known to be extremely volatile. For the 2011-2015 period in Victoria the mean price per megawatt hour (mWh) was $42.06, with 99\% of oberservations less than \$100/mWh, but on 39 occaisions the price spiked greater than \$1,000/mWh, reaching as high as \$9,974/mWh on November 29 2012. For this reason market participants are increasingly interested in a predictive density forecast, as the probability that the price will rise above a particular threshold in the next time period is more important than the mean value. To calculate a suitably accurate predictive density the Bayesian methodology is to be adopted in this thesis, as it can average over parameter and model uncertainty that is often ignored in frequentist methods. See \citet{Geweke2006} for a review on Bayesian forecasting and \citet{Gneiting2014} for density forecasting in general.

@@

Alternatives to MCMC include Variational Bayes (VB) and Expectation Propogation (EP), which aim to substitute the true posterior distribution for a vector of unknown parameters $\theta$, denoted by $p(\theta | y)$, with some approximate distribution denoted by $q(\theta | \lambda)$. Here $\lambda$ is a set of auxillary parameters associated with the approximation and not the model. For example, if $q(\theta | \lambda)$ is Gaussian $\lambda$ would be the mean and variance that parameterise $q$. Variational Bayes will be explored as an alternative computation strategy to MCMC in the context of short term electricity load forecasting using data from the 2011-2015 period in Victoria. The advantages in computation time for VB will be compared to the loss of statisitcal accuracy associated with the use of an approximation of the posterior instead of the truth.

@@

\begin{figure}[h]
\label{MCMCplot}
\centering
\includegraphics[scale = 0.5]{MCMC.png}
\caption{MCMC draws for the AR(2) model with $T = 150, \phi_1 = 0.75, \phi_2 = 0.2$ and $\sigma^2 = 1$. Approximately 24\% of both $\phi$ parameters were accepted by the Metropolis Hastings step, then the first 10,000 draws were discarded and the remainder thinned to 1 in 200 to make dependence between draws negligible.}
\end{figure}

@@

Restricting the approximating distribution to a factorisable family may be unsatisfactory in many applications where the dependence between parameters is important, as in the AR(2) model where Figure (\ref{MCMCplot}) shows strong dependence between $\phi_1$ and $\phi_2$. Without factorisation the simple method to maximise the ELBO as in MFVB is unavailable. \citet{Paisley2012} and \citet{Ranganath2014} have adapted a gradient ascent algorithm for use in Variational Bayes, resulting in what is known as Stochastic Variational Bayes (SVB).

@@

The literature for electricity forecasting is wide, and contains models based on many different fields, such as game theory, time-series modelling and neural networks, see \citet{Weron2014} for a recent review. Density forecasting for short term electricity load has had less attention, with \citet{Fan2012} and \citet{He2016} being notable examples, which used boot strapping and quantile regression respectively to generate forecast densities. Both of these approaches create forecast densities that are conditioned on the model and estimated parameters being correct, while the Bayesian approach used in this thesis can average over parameter and model uncertainty leading to a more accurate forecast density. This thesis also aims to re-estimate the model after each data point is observed to fully take advantage of any extra information on changes in parameters and thus further increase the accuracy of the forecast density.
